### 2011-03-09 Multiresolution_framework/ltx/ResearchHighlights.tex
added research highlights

%  Created by Parham Aram on 09-03-2011
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[]{article}

\begin{document}

\textbf{Research Highlights}
\begin{itemize}
	\item We derive a model-based, data-driven estimator of continuum neural field parameters.
	\item Both macroscopic and microscopic behavior of the system can be represented simultaneously.
	\item We show how to infer intracortical connectivity kernel with an arbitrary shape from data.
\end{itemize}

\end{document}