### 2011-03-12 Multiresolution_framework/ltx/ResearchHighlights.tex
edit to research highlights

%  Created by Parham Aram on 09-03-2011
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[]{article}

\begin{document}

\textbf{Research Highlights}
\begin{itemize}
	\item We derive a multi-resolution model-based, data-driven estimator of continuum neural field parameters.
	\item With the multi-resolution approach both the macroscopic and microscopic dynamics of the system can be represented simultaneously.
	\item We show for the first time how to infer an arbitrary shaped intracortical connectivity kernel from data.
\end{itemize}

\end{document}