### 2012-06-17 Multiresolution_framework/ltx/SubmissionLetter.tex
Paper submitted to NeuroImage.

%
%  untitled
%
%  Created by Dean Freestone on 2010-06-28.
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[a4paper,10pt]{letter}
\usepackage{anysize} 									%<-use this for checking equation length

\marginsize {1.7cm}{1.5cm}{1cm}{1cm} 				%<-use this for checking equation length

\signature{P. Aram}
\address{Theoretical Neuroscience Group \\
UMR 1106\\ Institut de Neurosciences des Systemes \\ 13385 Marseille \\ France}

\begin{document}
\vspace{-1cm}
\begin{letter}{Editor/Reviewer \\ Neuroimage}

\opening{Dear Sir or Madam,}

% \emph{Please explain why this manuscript is suitable for publication in PLoS Computational Biology; why will your paper inspire the other members of your field, and how will it drive research forward?}

I am writing to seek publication of our research article titled `Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model' in Neuroimage.

The novel aspects of this paper include:
\begin{itemize}
	\item a multi-resolution model-based, data-driven estimator of continuum neural field parameters ;
	\item simultaneous representation of both the macroscopic and microscopic dynamics of the system ;
	\item a derivation of an algorithm for estimating intracortical connectivity with an arbitrary shape using the Amari type neural field equations.
\end{itemize}

Thank you for your time and consideration in reviewing this article. We look forward to hearing the outcome of the review.

\closing{Best Regards,}

\end{letter}
\end{document}