### 2012-08-23 Multiresolution_framework/ltx/ResubmissionLetter.tex
Revised version submitted to NeuroImage

%
%  untitled
%
%  Created by Dean Freestone on 2010-06-28.
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[a4paper,10pt]{letter}
\usepackage{anysize} 									%<-use this for checking equation length

\marginsize {1.7cm}{1.5cm}{1cm}{1cm} 				%<-use this for checking equation length

\signature{P. Aram}
\address{Theoretical Neuroscience Group \\
UMR 1106\\ Institut de Neurosciences des Systemes \\ 13385 Marseille \\ France}

\begin{document}
\vspace{-1cm}
\begin{letter}{Editor/Reviewer \\ Neuroimage}

\opening{Dear Sir or Madam,}

    We thank the reviewers for their encouraging comments and their contribution to improving this paper. Please find below our responses to the reviewers questions and suggestions regarding the manuscript. 
\\

Thank you for your time and consideration in reviewing this article. We look forward to hearing the outcome of the review.

\closing{Best Regards,}

\end{letter}
\end{document}